---
title: "Simulation study comparing rule-based and model-based adaptive designs"
author: "James Watson"
date: "October 10, 2019"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE, 
                      include = TRUE, cache.lazy = F,
                      fig.width = 9, fig.height = 9,
                      fig.pos = 'H', dev = 'png', dpi = 300)
library(tictoc)
library(truncnorm)
library(rootSolve)
library(RColorBrewer)
library(boot)
library(rstan) 
library(doParallel)
library(foreach)
FORCE_RERUN = F
N_cores = 8

source('dose_response_model_stan.R') # compile stan dose-response model
```

## Setup model parameters

```{r}
source('Core_functions.R')
source('Simulation_functions.R')
source('plotting_functions.R')
source('wrapper_function_all_scenarios.R')
TEL = 0.95  # Target Efficacy Level
MTT = 0.05  # Maximum Tolerated Toxicity
N_max = 260 # Maximum number of patients recruited
max_increment = 1 # Maximum dose increment to doses previously unseen
N_trials = 2000
N_batch = 4
Randomisation_p_SOC = 0.2 # proportion randomised to the standard of care dose
starting_dose = 12 # starting dose in adaptive arm
SoC = 8 # Standard of Care

# function that solves for the beta value based on interpretable parameters
solve_beta = function(alpha_val, v_star, y_star){
  beta_val = ( logit(y_star) - alpha_val ) / log2(v_star)
  return(beta_val)
}

optimal_doses = array(NA,dim = 7)
```

## Priors

```{r prior}
#******* Prior point estimates *********
Prior_MTD = 32;                    # prior estimate of the Maximum Tolerated Dose
Prior_alpha_tox = logit(1/1000)    # prior estimate of the toxicity after 1 vial
Prior_beta_tox = solve_beta(alpha_val = Prior_alpha_tox, v_star = Prior_MTD, y_star = MTT)

#******* Prior uncertainty estimates *******
prior_model_params = list(beta_tox = Prior_beta_tox,
                          beta_tox_sd = .05,
                          alpha_tox=Prior_alpha_tox,
                          alpha_tox_sd = 2,
                          mu_antivenom = 8,
                          mu_antivenom_sd = 3,
                          sd_antivenom = 5,
                          sd_antivenom_sd = 2)
```



## Simulation 1

```{r simulation_scenario_1, cap="Simulation scenario 1"}
true_TED = 20  # 
true_antivenom_mean = 10
true_antivenom_sd = (true_TED - true_antivenom_mean)/qnorm(.95)

true_alpha_tox = logit(1/500)  # toxicity at 1 vial
true_MTD = 8  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)

# well-specified simulation truth
model_params_true = list(beta_tox = true_beta_tox,
                         alpha_tox=true_alpha_tox,
                         mu_antivenom=true_antivenom_mean, 
                         sd_antivenom=true_antivenom_sd)
optimal_doses[1] = min(true_MTD, true_TED)
tic()
Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 1',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 1',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()


compare_rule_vs_model(sim_title = 'Simulation scenario 1',true_model = NULL,
                      model_params_true = model_params_true,
                      prior_model_params = prior_model_params,use_SoC_data = T,
                      idiosyncratic = F,plot_vstar = T, MTT = MTT, TEL = TEL)
```


## Simulation 2

```{r simulation_scenario_2}
true_TED = 8  
true_antivenom_mean = 4
true_antivenom_sd = (true_TED - true_antivenom_mean)/qnorm(.95)

true_alpha_tox = logit(1/500)  # toxicity at 1 vial
true_MTD = 20  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)


# well-specified simulation truth
model_params_true = list(beta_tox = true_beta_tox,
                         alpha_tox=true_alpha_tox,
                         mu_antivenom=true_antivenom_mean, 
                         sd_antivenom=true_antivenom_sd)
optimal_doses[2] = min(true_MTD, true_TED)

tic()
Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 2',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 2',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()

compare_rule_vs_model(sim_title = 'Simulation scenario 2',true_model = NULL,
                      model_params_true = model_params_true,
                      prior_model_params = prior_model_params,use_SoC_data = T,
                      idiosyncratic = F,plot_vstar = T, MTT = MTT, TEL = TEL)
```


## Simulation 3

```{r simulation_scenario_3}
true_TED = 60  
true_antivenom_mean = 30
true_antivenom_sd = (true_TED - true_antivenom_mean)/qnorm(.95)

true_alpha_tox = logit(1/1000)  # toxicity at 1 vial
true_MTD = 30  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)

# well-specified simulation truth
model_params_true = list(beta_tox = true_beta_tox,
                         alpha_tox = true_alpha_tox,
                         mu_antivenom=true_antivenom_mean, 
                         sd_antivenom=true_antivenom_sd)

optimal_doses[3] = min(true_MTD, true_TED)

tic()
Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 3',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 3',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()

compare_rule_vs_model(sim_title = 'Simulation scenario 3',true_model = NULL,
                      model_params_true = model_params_true,
                      prior_model_params = prior_model_params,use_SoC_data = T,
                      idiosyncratic = F,plot_vstar = T, MTT = MTT, TEL = TEL)
```

## Simulation 4

```{r simulation_scenario_4}
true_TED = 30  
true_antivenom_mean = 15
true_antivenom_sd = (true_TED - true_antivenom_mean)/qnorm(.95)

true_alpha_tox = logit(1/1000)  # toxicity at 1 vial
true_MTD = 60  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)

# well-specified simulation truth
model_params_true = list(beta_tox = true_beta_tox,
                         alpha_tox = true_alpha_tox,
                         mu_antivenom=true_antivenom_mean, 
                         sd_antivenom=true_antivenom_sd)
optimal_doses[4] = min(true_MTD, true_TED)

tic()
Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 4',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = model_params_true,
                true_model = NULL,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 4',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()

compare_rule_vs_model(sim_title = 'Simulation scenario 4',true_model = NULL,
                      model_params_true = model_params_true,
                      prior_model_params = prior_model_params,
                      use_SoC_data = T, idiosyncratic = F,plot_vstar = T,
                      MTT = MTT, TEL = TEL)
```


# Simulation 5 
Idiosyncratic toxicity

This simulation assumes a dose-independent probability of having a toxic event. 
We simulate a trial whereby the overall toxicity is 15%

```{r idiosyncratic}
true_model = list(tox = approxfun(x = c(1,10), y = c(.15,.15), rule = 2),
                  eff = approxfun(x = 0:100, 
                                  y = pnorm(q = 0:100, mean = true_antivenom_mean, 
                                            sd = true_antivenom_sd),rule=2))
optimal_doses[5] = 0

tic()
# Model based simulation
Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 5',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,use_SoC_data = T)


# Rule based simulation
Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = 10*N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 5',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC, use_SoC_data = T)
toc()

# # Comparison in mis-specified idiosyncratic case
compare_rule_vs_model(sim_title = 'Simulation scenario 5',
                      model_params_true = NULL,
                      true_model = true_model,
                      prior_model_params = prior_model_params,
                      use_SoC_data = T, idiosyncratic = T,plot_vstar = F, 
                      MTT = MTT, TEL = TEL)
```


## Simulation 6

Same as simulation 2 but with mis-specified efficacy dose response

```{r simulation_scenario_6}
true_alpha_tox = logit(1/500)  # toxicity at 1 vial
true_MTD = 20  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)

true_rate = 1/2.7
# mis-specified simulation truth
true_model = list(eff = approxfun(x =0:100, y = pexp(q = 0:100, rate = true_rate), rule = 2),
                  tox = approxfun(x = 0:200, 
                                  y = inv.logit(true_alpha_tox + true_beta_tox*log2(0:200)),rule=2))
optimal_doses[6] = qexp(p = TEL, rate = true_rate)

tic()
Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 6',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 6',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()

compare_rule_vs_model(sim_title = 'Simulation scenario 6',true_model = true_model,
                      model_params_true = NULL,
                      prior_model_params = prior_model_params,use_SoC_data = T,
                      idiosyncratic = F,plot_vstar = T,MTT = MTT, TEL = TEL)
```

## Simulation 7

Same as simulation 4 except with exponentially distributed venom volume

```{r simulation_scenario_7}
true_alpha_tox = logit(1/1000)  # toxicity at 1 vial
true_MTD = 60  # simulation truth for the MTD 
true_beta_tox = solve_beta(alpha_val = true_alpha_tox, v_star = true_MTD, y_star = MTT)

true_rate = 1/10
# mis-specified simulation truth
true_model = list(eff = approxfun(x =0:200, y = pexp(q = 0:200, rate = true_rate), rule = 2),
                  tox = approxfun(x = 0:500, 
                                  y = inv.logit(true_alpha_tox + true_beta_tox*log2(0:500)),rule=2))
optimal_doses[7] = qexp(p = TEL, rate = true_rate)

tic()
Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 7',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'model_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)

Full_Simulation(model_params_true = NULL,
                true_model = true_model,
                prior_model_params = prior_model_params,
                N_trials = N_trials,
                MTT = MTT, TEL = TEL,
                N_max = N_max,
                N_batch = N_batch,
                max_increment = max_increment,
                Randomisation_p_SOC = Randomisation_p_SOC,
                sim_title = 'Simulation scenario 7',
                FORCE_RERUN=FORCE_RERUN,
                N_cores = N_cores,
                design_type = 'rule_based',
                starting_dose = starting_dose,
                SoC = SoC,
                use_SoC_data = T)
toc()

compare_rule_vs_model(sim_title = 'Simulation scenario 7',true_model = true_model,
                      model_params_true = NULL,
                      prior_model_params = prior_model_params,
                      use_SoC_data = T, idiosyncratic = F,plot_vstar = T,
                      MTT = MTT, TEL = TEL)
```

# Summary of all trials

```{r summary_sim_studies}
specification_type = 'well_specified'
mypallete = brewer.pal(9,'Set1')[c(1,2,7)]
par(las=1,bty='n',family = 'serif', cex.lab=1.5, cex.axis=1.5)
layout(mat = matrix(data = c(1,1,1,2,2,2,3,3,3,4,4,4,5,5,6,6,7,7),nrow = 3,byrow = T))
for(ss in 1:7){
  if(ss>4) specification_type = 'mis_specified'
  f_name_rule = paste('SimulationOutputs/', 'Simulation scenario ', ss, 
                      '_', 'rule_based', '_', specification_type, 
                      '_', 'All_data', '.RData', sep = '')
  f_name_model = paste('SimulationOutputs/','Simulation scenario ', ss,
                       '_','model_based', '_', specification_type, 
                       '_', 'All_data', '.RData', sep = '')
  load(f_name_model)
  Summary_model = Summary_trials
  load(f_name_rule)
  Summary_rule = Summary_trials
  rm(Summary_trials)
  
  Summary_model_opt = Summary_model[, grep('assigned_dose', colnames(Summary_model))]
  opt_dose = optimal_doses[ss]
  
  Summary_model_opt_qs = 10* (apply(Summary_model_opt,1,quantile,probs = c(0.025,0.975),na.rm=T) - opt_dose)
  Summary_rule_opt_qs = 10 * (apply(Summary_rule,1,quantile,probs = c(0.025,0.975),na.rm=T) - opt_dose)
  
  plot(10*(rowMeans(Summary_model_opt,na.rm=T) - opt_dose), type='l', lwd=3, 
       ylim=range(c(0,Summary_model_opt_qs)),col=mypallete[2],
       xlab='Number of patients enrolled', 
       ylab = 'Difference from optimal (mL)')
  polygon(x = c(1:nrow(Summary_model),rev(1:nrow(Summary_model))), 
          y = c(Summary_model_opt_qs[1,], rev(Summary_model_opt_qs[2,])), 
          col = adjustcolor(mypallete[2],alpha.f = .2), border = NA)
 
  lines(10*(rowMeans(Summary_rule,na.rm=T) - opt_dose), lwd=3, lty=2,col=mypallete[3])
  polygon(x = c(1:nrow(Summary_model),rev(1:nrow(Summary_model))), 
          y = c(Summary_rule_opt_qs[1,], rev(Summary_rule_opt_qs[2,])), 
          col = adjustcolor(mypallete[3],alpha.f = .2), border = NA)
  mtext(text = ss,side = 3,line = 1,adj = 0)
  abline(h=0, col=1)
  
  if(ss==1){
    legend('topright',col = mypallete[2:3], legend = c('Model based', 'Rule based'),
           cex=1.2,lwd=3,lty=1:2)
  }
}
```

